Adaptive Sequential Model Selection

### 2015-06-17 paper.tex
some more work, about to do major reorg

Define the {\em completion index} $k_0(Y) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model. By construction, $F\in M_k \iff k \geq k_0$. Our goal is to examine the data and output a {\em stopping rule}, i.e. an estimator $\hk(Y)$ of $k_0$. We consider $M_k$ to be ``rejected'' if $k < \hat k$, and ``accepted'' otherwise, so that $V=(\hat k - k_0)_+$ is the number of mistaken rejections.

@@

\begin{comment}
As we will see, {\em selected-model} tests can be dramatically more powerful than {\em saturated-model} tests at early steps in the model path when most of the signal variables have not yet entered the model.

Under weak conditions, selected-model $p$-values are independent. By contrast, saturated-model $p$-values generically produce non-independent null $p$-values.

We could have instead defined the ``variable-centric'' type I error $\tV$ as the number of noise variables incorrectly included in the final model, with $\tR$ denoting the total number of variables included. Here, by ``noise variable'' we mean one that has a non-zero coefficient in the full model. Using selective inference to control the variable-wise FWER $\P_\mu(\tV>0)$ and FDR $\E[\tV / (\tR \vee 1)]$ is an interesting topic for further study.

There is some ambiguity involved in deciding how to generalize $z$- and $t$-tests to the selective case. For example, \citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$ in the case of linear regression under the upper model $M_\infty:\; Y\sim \cN(\mu, \sigma^2 I)$:
\begin{description}
\item[Complete Null:] The previous model, $M_{k-1}$, is correct; i.e., that $\mu = X_{E_{k-1}}\beta$ for some $\beta$.
\item[Incremental Null:] $M_{k-1}$ may not be correct, but $M_k$ is no improvement: i.e., that $X_EX_E^\dagger\mu = \proj_{E_{k-1}}\mu$
\end{description}
. \citet{fithian2014optimal} discusses several different tests

\WFcomment{Also say: Controlling for more variables could take us farther from a causal effect; furthermore, coefficient in larger model is much harder to interpret.}
\end{comment}

@@

Having discussed methods for constructing single-step $p$-values, we turn now to the problem of constructing a {\em stopping rule}; that is, an estimator $\hk$ of the completion index $k_0(Y)\in \{0,\ldots,d,\infty\}$. $M_k$ is ``rejected'' if and only if $\hk>k$. Thus $\hk$ is the number of models we rejected, while $k_0$ is the number that we should have rejected. The number of type I errors, then, is $(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.

\subsection{Stopping Rules}

We will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values, which have been studied in the literature by \WFcomment{literature review here.} We will focus on three such stopping rules: Basic Stop, proposed by \WFcomment{whom?}, and the more refined Strong Stop and Forward Stop, proposed by \citet{gsell2013sequential}.

While Strong Stop and Forward Stop tend to give more powerful stopping rules, they require independence among the $p$-values. Sections~\ref{sec:pValsIndep}--\ref{sec:modelSSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent. We will see that, typically, selected-model $p$-values are sequentially independent while saturated-model $p$-values are not.

\subsubsection{Basic Stop:}

\WFcomment{Is there a better name for this? Can we find it in the literature?}

The most obvious sequential procedure is to reject at each step until the first time $p_k > \alpha$, which we can formalize as
\[
\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
  p_k > \alpha\right\} - 1
\]
We will call this procedure {\em Basic Stop}.

If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
\[
\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha.
\]
\WFcomment{Trivial, but no doubt someone else proved this first...}

@@

\subsubsection{Strong Stop:}

\citet{gsell2013sequential} propose another stopping rule that controls the model-wise FWER, which they called {\em strong stop}. Define
\[
  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
    \leq \frac{\alpha k}{d}\right\}
\]
Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, strong stop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.

@@

For random $k_0$, it is easy to see that the $p$-values simply need to satisfy~\eqref{eq:indepCond_fixed_k0} conditional on the value of $k_0$:
\begin{equation}\label{eq:indepCond_random_k0}
  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{i=k+1}^d \alpha_i
\end{equation}
If~\eqref{eq:indepCond_random_k0} holds, then strong stop controls theprobability of $V>0$ conditionally on $k_0$, and thus also marginally.

\subsubsection{Forward Stop:}

\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the model-wise FDR:
\[
  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
\]
If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
\[
\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
\]
is a Gamma random variable with mean $V/k$ and variance $V/k^2$. We stop the last time the estimate of FDR is less than $\alpha$. Forward stop requires the same condition independence condition~\eqref{eq:indepCond_random_k0} as strong stop does.

>>>>

Define the {\em completion index} $k_0(Y) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model. By construction, $F\in M_k \iff k \geq k_0$. Our goal is to examine the data and output a {\em stopping rule}, i.e. an estimator $\hk(Y)$ of $k_0$. We consider $M_k$ to be ``rejected'' if $k < \hat k$, and ``accepted'' otherwise. Thus $\hk$ is the number of models we rejected, while $k_0$ is the number that we should have rejected. The number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.

@@

@@

Having discussed methods for constructing single-step $p$-values, we turn now to the question of using them to construct a stopping rule $\hat k$. Specifically, we will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values and must always reject an initial block of corresponding hypotheses.

Section~\ref{sec:randomHyp} outlines several proposals for ordered testing of hypotheses and discusses what error rates they control and under what conditions. We focus on proposals from \citet{gsell2013sequential} and \citet{li2015accumulation}.

These two papers both assume that the hypotheses $H_1, \ldots, H_d$ are fixed, and in particular that $k_0$ is fixed. In our problem, the model path is random and thus so is $k_0$. As we will see in Section~\ref{sec:randomHyp}, \WFcomment{... that this is not a problem}.

While StrongStop, ForwardStop, and the accumulation tests tend to give more powerful stopping rules, they require independence among the $p$-values. Sections~\ref{sec:pValsIndep}--\ref{sec:modelSSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent. We will see that, typically, selected-model $p$-values are sequentially independent while saturated-model $p$-values are not.

\subsection{Ordered Testing of Random Hypotheses}\label{sec:randomHyp}

\subsubsection{BasicStop}
The most obvious procedure is simply to reject at each step until the first time that $p_k > \alpha$, which we can formalize as
\[
\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
  p_k > \alpha\right\} - 1
\]
We will call this procedure {\em BasicStop}. \WFcomment{Is there a better name for this? Can we find it in the literature?}

If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
\[
\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha.
\]

@@

\subsubsection{StrongStop}

\citet{gsell2013sequential} propose another stopping rule that controls the model-wise FWER, which they called {\em strong stop}. Define
\[
  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
    \leq \frac{\alpha k}{d}\right\}
\]
Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, StrongStop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.

@@

For random $k_0$, it is easy to see that the $p$-values simply need to satisfy~\eqref{eq:indepCond_fixed_k0} conditional on the value of $k_0$:
\begin{equation}\label{eq:indepCond_random_k0}
  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{i=k+1}^d \alpha_i
\end{equation}
If~\eqref{eq:indepCond_random_k0} holds, then strong stop controls the probability of $V>0$ conditionally on $k_0$, and thus also marginally.

\subsubsection{ForwardStop}

\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the model-wise FDR:
\[
  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
\]
If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
\[
\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
\]
is a Gamma random variable with mean $V/k$ and variance $V/k^2$. We stop the last time the estimate of FDR is less than $\alpha$. Forward stop requires the same conditional independence condition~\eqref{eq:indepCond_random_k0} as strong stop does.

\subsubsection{Accumulation Tests and HingeExp}

\WFcomment{Discuss Rina's stuff.}
